Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 schemes.tex

In this section, we present a description of the considered coding schemes
measured with the \ac{Raspi} testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations that provide
better processing speeds, particularly relevant for the \ac{Raspi}.
Finally, we include a review of the \todot{this sentence is incomplete}

@@

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding, all
the original data packets $P_j, j \in [1,g]$, and each of $B$ bytes, are used to
create coded packets. In the following subsections, we describe the basic
functionalities of \ac{RLNC}.

@@

For indicating a receiver which packets were utilized
to create a coded one, a simple yet versatile choice is to append
its coding coefficients as a header of the coded packet. Hence, an
amount of overhead is included in every coded packet given that we
need to provide some signaling needed for decoding. The coding
coefficients overhead amount for packet $i$, $|v_i|$, can be quantified
as:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{ij}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
\label{sssec:decoding}
To be able to decode a batch of $g$ packets, a \ac{l.i.} set of $g$
coded packets, $C_i,\ i \in [1,g]$ is required at a decoder.
Once this set has been collected for a decoder,
the original packets can be found by computing the solution of a system
of linear equations using \ac{GF} aritmethics. Thus, we define
$\textbf{C} = \left[C_1 \ldots C_g \right]^T$,
$\textbf{P} = \left[P_1 \ldots P_g \right]^T$ and the coding matrix
$\textbf{V}$ that collects the coding coefficients for each of the $g$
coded packets, as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{11} & \ldots & v_{1g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g1} & \ldots & v_{gg} \\
\end{array}
\right]
\end{align}
%
Algebraically, decoding simplifies to finding the inverse of
$\textbf{V}$ in the linear system
$\textbf{C} = \textbf{V} \textbf{P}$, which can achieved using
efficient Gaussian elimination techniques \cite{fragouli2006network}.
On real applications, decoding is performed on-the-fly, e.g., the pivots are
computed as packets are progressively received, in order to minimize the
computation delay for each step.

@@

\subsubsection{Recoding}
As an inherent property of \ac{RLNC}, an intermediate node in the network
is able to create new coded packets without needing to decode previously
received ones from an encoding source. Therefore, \ac{RLNC} is an end-to-end
coding scheme that permits to \textit{recode} former coded packets at any
point in the network without inquiring in local decoding processing. In
principle, a recoded packet should be indistinguishable from a coded one.
Thus, we define a recoded packet as $R_i$ and consider the coding
coefficients $w_{i1}, \ldots, w_{ig}$ as the used ones to
create $R_i$. Later, a recoded packet can be written as:
%
\begin{align}
\label{eq:recoded_packet}
R_i = \bigoplus_{j=1}^{g} w_{ij} \otimes C_j  ,\ \forall i \in [1,\ldots)
\end{align}

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously coded
packet $C_j$. These coding coefficients are again uniformly and randomly chosen
from $GF(q)$. Still, each of the $w_{ij}$ are not the appended recoding
coefficients for signalling since the they will only retrieve previously
coded packets. The analytical expression for the appended coefficients
after recoding is obtained by calculating recoded packets as function of the
original packets. Therefore, we define the local coding matrix $\textbf{W}$
in the same way as it was made for $\textbf{V}$. Thus, the local coding matrix
can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right]
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g]
\end{align}
%
This procedure is carried by all the recoders in a given network therefore
allowing any decoder to compute the original data after Gaussian elimination,
regardless of the amount of times where recoding was performed. Similarly
as with the encoding operation, any decoder that collects a set of
$g$ \ac{l.i.} recoded packets with their respective coefficients, will
be able to decode the data as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a coded packet
as in \ac{RLNC}, an encoder sets more coding coefficients to zero when
generating a coded packet with the purpose of reducing the overall processing.
Decoding is the same as in \ac{RLNC}, but given that the coding matrices are now
sparse, we will have less operations performed in the decoding process.
Recoding, although theoretically possible, is omitted in practical systems since
it is too complex to keep packets sparse after recoding
[\textbf{reference}]\todot{add ref}. In what follows, we describe the coding
scheme with a basic method to produce sparse coded packets and later we indicate
a second one, giving reference to their implementation performance.

\subsubsection{Method 1: Fixing the Coding Density}
A way to control the amount of non-zero coding coefficients, is to set a
fixed ratio of non-zero coefficients in the encoding vector of size $g$. We
refer to this fixed ratio as the average coding density $d$. Thus, for any
coded packet $C_i$ with coding coefficients $v_{ij},\ j \in [1,g]$, its
coding density is defined as follows:
%
\begin{align}
\label{eq:coding_density}
d = \frac{\sum_{j=1}^{g} f(v_{ij})}{g},\ \
    f(v_{ij}) =
    \begin{cases}
        0 &,\ v_{ij} = 0 \\
        1 &,\ v_{ij} \neq 0
    \end{cases}
\end{align}
%
From the density definition in \eqref{eq:coding_density}, it can be
observed that $0 \leq d \leq 1$. As $g$ increased, we obtain more
granularity in the density. Notice that the special case $d = 0$ has no
practical purpose since it implies just zero padded data. So, a practical
range for the density excludes the zero case, $0 < d \leq 1$. Moreover,
it can be also observed that the case of $d = 1$ reduces to the \ac{RLNC}
coding scheme described in Section \ref{ssec:RLNC}.

To achieve a desired average coding density in the coefficients, for each
of them, we utilize a set of Bernoulli random variables all with parameter
$d$ as it success probability, i.e. $\mathbb{B}_j \sim Bernoulli(d)\ ,\ \forall
j \in [1,g]$. In this way, we can represent a coded packet in \ac{SRLNC} as:
%
\begin{align} \label{eq:sparse_coded_packet}
    C_i  &= \bigoplus_{j=1}^{g} \mathbb{B}_j v_{ij} \otimes P_j ,\ v_{ij}
    \neq 0\
    \forall i \in [1,\ldots) ,\ \
    d \in
    \begin{cases}
        {(0,0.5]} & ,\ q = 2 \\
        {(0,1]}   & ,\ q > 2
    \end{cases}
\end{align}
%
In \eqref{eq:sparse_coded_packet}, we have the requirement for the coding
coefficient to not be zero, since we want to ensure a coding coefficient
is generated for any random trial where $\mathbb{B}_j = 1$. Therefore, in our
implementation of \ac{SRLNC} we exclude the zero element and then pick
uniformly distributed random elements from $GF(q)-\{ 0 \}$. Also, we have
specified a dependency on the field size for practical
density values. In case of employing $GF(2)$, the
maximum plausible density is restricted up to 0.5 since higher values incur
in more frequent \ac{l.d.} coded packets \textbf{Include reference here}.

@@

The drawback of this scheme is that coded packets from the encoder
become more \ac{l.d.} on each other as the density is reduced.
This leads to transmission overhead since another coded packet
is required to be sent for every reception of a redundant packet.
Also, particularly with this method, we may still generate a coded packet
which does not contain any data at all. For example, we might find
the case where $\left[\mathbb{B}_1, \ldots, \mathbb{B}_g \right] =
\mathbf{0}$. In that case, the encoder discards the coded packet
and tries to generate a new one. This situation occurs more frequently
as the density is reduced.

\subsubsection{Method 2: Sampling the Amount of Packets to Combine}
The method described from \eqref{eq:sparse_coded_packet} permits
to obtain a fast implementation in terms of execution time for
$d \geq 0.3$ \textbf{Include reference here}. Nevertheless, for $d < 0.3$
the implementation becomes slow given that more Bernoulli trials are
required in total for all random variables to generate a coded packet.
Thus, we introduce a second method that permits a fast implementation
for low coding densities \textbf{Same reference here}.

For this method, we first obtain the amount of packets
that we will combine to create a coded packet. To do it so, a random
number $\mathbb{N}$ of the original packets is used to produce a
coded packet. For our case, $\mathbb{N}$ is binomially distributed
with parameters $g$ for the number of trials and $d$ for its success
probability, e.g. $\mathbb{N} \sim Binomial(g,d)$. However,
when sampling from this distribution, the case
$\mathbb{N} = 0$ occurs with a non-zero probability. In order to handle
this special case, our implementation considers
$\mathbb{K} = \max(1,\mathbb{N})$ as the final amount of packets to be
mixed together. In this way, we always ensure that at least one packet is
encoded.

The only caveat is that the case of $\mathbb{K} = 1$ occurs slightly
more often than $\mathbb{N} = 1$ in the original distribution, but for
the considered density range in this method, this is not a significant
modification \textbf{Same reference here}. Then, once the distribution for the
number of packets to mix has been defined, we sample a value $n$ from
$\mathbb{N}$ and compute $k = \max(1,n)$. Later, we create a create a
set $\mathcal{K}$ with cardinality $k$, e.g. $|\mathcal{K}| = k$,
where the elements in $\mathcal{K}$ are the indexes of the packets that are
going to be considered for making a coded packet. To compute the indexes
of the set $\mathcal{K}$, we do the following algorithm in pseudo-code: \\
\ \\
\begin{algorithm}[H]
 \label{alg:k_set}
 \KwData{$k$: Size of $\mathcal{K}$. $g$: Generation Size}
 \KwResult{$\mathcal{K}$: The set of non-repeated indexes}
 $\mathcal{K}$ = \{ \}\;
 \While{$|\mathcal{K}| \neq k$ }{
  $i$ = Sample from $\mathbb{U}(1,g)$\;
  \If{$i \notin \mathcal{K}$}{
   Insert $i$ in $\mathcal{K}$\;
   }
 }
 \caption{Computation of the set of indexes for packet combination in SRLNC.}
\end{algorithm}
\ \\
Where in Algorithm~\ref{alg:k_set}, the notation $\mathbb{U}(1,g)$ stands
for a uniform discrete random variable with limits $1$ and $g$. The
pseudo-code in Algorithm~\ref{alg:k_set} indicates that the set $\mathbb{K}$
is filled with non-repeated elements taken uniformly at random from
$[1,\ldots,g]$. Finally, once the set of indexes has been defined, a coded
packet using \ac{SRLNC} can be written as shown in
\eqref{eq:sparse_coded_packet2}:
%
\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{m \in \mathcal{K}} v_{im} \otimes P_{m},\ v_{im} \neq 0,\ \forall i \in [1,\ldots)
    %L \subseteq \{1,\dots,g\}: |L|=k ,\
\end{align}

@@

Therefore, in \ac{TSNC}, a simple procedure for controlling this probability
is to gradually increase the coding density as the \ac{dof}\footnote{We refer
as degrees of freedom to the dimension of the linear span from the coded
packets that a decoder has at a given point} increases, e.g. as the
cardinality of the set of \ac{l.i.} coded packets increases. This enables
\ac{TSNC} to significantly reduce the complexity, particularly in the
beginning of the transmission process and also to control the decoding
delay. For \ac{TSNC}, we define the budget, $b \geq g$, to be a target
number of coded packets a transmitter wants to send to a receiver
for decoding a generation of $g$ packets. In some scenarios, we may set
$e$ defined as the excess of \ac{l.d.} packets sent from the encoder.
This helps to also define the budget as $b = g + e$.

The difference between $b$ and $g$ is equal to the losses in the channel and the
losses due to linear dependencies. Therefore, in a lossless channel, the budget
is
\begin{equation}\label{eq:budget}
    b(g,d)=\sum^{g-1}_{i=0}\frac{1}{P(i,g,d)}
\end{equation}

where $P(i,g,d)$ is the probability of receiving an innovative coded packet
after receiving $i$ \ac{l.i.} packets with a coding density $d$. In our implementations, we considered the the upper bound for the innovation probability from \cite{feizi2014tunable}, given as
\begin{equation}
    P(i,g,d)\geq 1-(1-d)^{g-i}
\end{equation}

Provided a desired budget $b$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k),\
k \in [1, \ldots)$ as the rank $r$ of the decoder\footnote{The rank of
a decoder is the rank of its coding matrix at a given point in the
transmissions process.} where a decoder has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in [1,\ldots)
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in [\textbf{reference}].\todot{@Chres this reference is missing}

The transmissions made between feedback packets are called the density regions
since the same estimated density is used for encoding sparse packets. Once new
feedback is received, an encoder recalculates the coding density for the new
region until the next feedback packet is received. before the coding density can
be estimated, it is essential that the encoder has a good estimate of the
decoder \ac{dof}, but also the remainder of the total budget. To calculate the
density, we use bisection to estimate a fixed density for the density region
that satisfies the budget for the $k$-th density region:
\begin{equation}
b(s(k-1),r(k),g,d) = \sum^{r(k)}_{i=r(k-1)} \frac{1}{P(i,g,d)} = \sum^{r(k)}_{i=r(k-1)} \frac{1}{1-(1-d)^{g-i}} = \frac{b}{2^k}, k\in [1,...).
\end{equation}
%
% Where the budget is half of the remaining budget for all the density regions
% except the last one, to which it is assigned the remaining of the budget.
The feedback scheme based on the rank reports of \eqref{eq:feedback_dofs}
roughly splits the remaining coding regions in two halves. In other
words, the first half is the region where the encoder currently operates. Once
an encoder finishes with this region, it proceeds with second half where again
it splits this into to two region and so on until it finishes on transmitting
the whole generation. This is also the case for the budget that is split equally
among the two new regions. The very last region will be assigned the full
remainder of the total budget and the coding density will not vary.

@@

When designing multi-threaded algorithms for network coding operations, it is
possible to implement the decoding by combining the matrix inversion and the
matrix multiplication, e.g. performing the Gauss-Jordan algorithm over the
coding matrix $\textbf{V}$ while performing, in parallel, row operations on the
coded data $\textbf{C}$. For example, in~\cite{5061951} and~\cite{4262451} the
parallelization of the row operations are optimized for \ac{GPU} and \ac{SMP}
systems respectively. However, the parallelization of such operations provide
limited speeds up for small block sizes ($\leq 2048$ bytes). The reason is that
to operate in a parallel fashion over the same coded packet $C_i$ requires
strained synchronization.

>>>>

In this section, we present a description of the considered coding
schemes measured with the \ac{Raspi} testbed. We introduce a
definition for the primitive coding operations, e.g. encoding,
decoding and recoding (where it applies) for each coding
scheme. Later, we address particular schemes which are obtained by
modifying the basic coding operations that provide better processing
speeds, which is particularly relevant for the \ac{Raspi}.  Finally,
we include a review of algorithms for network coding that exploit
multicore capabilities of the \ac{Raspi} 2.

@@

\ac{RLNC} is an example of intra-session \ac{NC}, i.e., data symbols
from a single flow are combined with each other. In this type of
network coding, all the original data packets $P_j, j \in [1,g]$, each
of $B$ bytes, are used to create coded packets using random linear
combinations of the original ones. In the following subsections, we
describe the basic functionalities of \ac{RLNC}.

@@

For indicating to a receiver how the packets were combined
to create a coded one, a simple yet versatile choice is to append
its coding coefficients as a header in the coded packet. Hence, an
amount of overhead is included in every coded packet given that we
need to provide some signaling needed for decoding. The coding
coefficients overhead amount for packet $i$, $|v_i|$, can be quantified
as:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{ij}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}].
\end{align}

\subsubsection{Decoding}
\label{sssec:decoding}
To be able to decode a batch of $g$ packets, a \ac{l.i.} set of $g$
coded packets, $C_i,\ i \in [1,g]$ is required at a decoder.
Once this set has been collected for a decoder,
the original packets can be found by computing the solution of a system
of linear equations using \ac{GF} aritmethics. Thus, we define
$\textbf{C} = \left[C_1 \ldots C_g \right]^T$,
$\textbf{P} = \left[P_1 \ldots P_g \right]^T$ and the coding matrix
$\textbf{V}$ that collects the coding coefficients for each of the $g$
coded packets, as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{11} & \ldots & v_{1g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g1} & \ldots & v_{gg} \\
\end{array}
\right].
\end{align}
%
Algebraically, decoding simplifies to finding the inverse of
$\textbf{V}$ in the linear system
$\textbf{C} = \textbf{V} \textbf{P}$, which can be achieved using
efficient Gaussian elimination techniques \cite{fragouli2006network}.
On real applications, decoding is performed on-the-fly, e.g., the
pivots are computed as packets are progressively received, in order to
minimize the computation delay for each step.

@@

\subsubsection{Recoding}
As an inherent property of \ac{RLNC}, an intermediate node in the network
is able to create new coded packets without needing to decode previously
received ones from an encoding source. Therefore, \ac{RLNC} is an end-to-end
coding scheme that permits to \textit{recode} former coded packets at any
point in the network without requiring a local decoding of the data. In
principle, a recoded packet should be indistinguishable from a coded one.
Thus, we define a recoded packet as $R_i$ and consider the coding
coefficients $w_{i1}, \ldots, w_{ig}$ as the used ones to
create $R_i$. Later, a recoded packet can be written as:
%
\begin{align}
\label{eq:recoded_packet}
R_i = \bigoplus_{j=1}^{g} w_{ij} \otimes C_j  ,\ \forall i \in [1,\ldots).
\end{align}

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously
coded packet $C_j$. These coding coefficients are again uniformly and
randomly chosen from $GF(q)$. However, these $w_{ij}$'s are not
appended to the previous coding coefficients.  Instead, the system
will update the previous coefficients. Due to the linearity of the
operation, this update reduces to recombining the coding coefficients
equivalent to each original packet with the weight of the
$w_{i,j}$.Therefore, we define the local coding matrix $\textbf{W}$ in
the same way as it was made for $\textbf{V}$. Thus, the local coding
matrix can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right].
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g].
\end{align}
%
This update procedure on the coding coefficients is carried by all the
recoders in a given network therefore allowing any decoder to compute
the original data after Gaussian elimination, regardless of the amount
of times recoding was performed and without incurring in any
additional overhead cost for signaling. Similar to the encoding
operation, any decoder that collects a set of $g$ \ac{l.i.} recoded
packets with their respective coefficients, will be able to decode the
data as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a
coded packet as in \ac{RLNC}, an encoder sets more coding coefficients
to zero when generating a coded packet with the purpose of reducing
the overall processing.  Decoding is the same as in \ac{RLNC}, but
given that the coding matrices are now sparse, we will have less
operations performed in the decoding process.  Recoding, although
theoretically possible, is omitted since it requires the use of
heuristics keep packets sparse after recoding, which is inherently
sub-optimal. In what follows, we describe the coding scheme with a
basic method to produce sparse coded packets and later we indicate a
second one, giving reference to their implementation performance.

\subsubsection{Method 1: Fixing the Coding Density}
A way to control the amount of non-zero coding coefficients, is to set a
fixed ratio of non-zero coefficients in the encoding vector of size $g$. We
refer to this fixed ratio as the average coding density $d$. Thus, for any
coded packet $C_i$ with coding coefficients $v_{ij},\ j \in [1,g]$, its
coding density is defined as follows:
%
\begin{align}
\label{eq:coding_density}
d = \frac{\sum_{j=1}^{g} f(v_{ij})}{g},\ \
    f(v_{ij}) =
    \begin{cases}
        0 &,\ v_{ij} = 0 \\
        1 &,\ v_{ij} \neq 0
    \end{cases}
\end{align}
%
From the density definition in~\eqref{eq:coding_density}, it can be
observed that $0 \leq d \leq 1$. As $g$ increased, we obtain more
granularity in the density. Notice that the special case of  $d = 0$ has no
practical purpose since it implies just zero padded data. So, a practical
range for the density excludes the zero case, $0 < d \leq 1$. Moreover,
it can be also observed that the case of $d = 1$ reduces to the \ac{RLNC}
coding scheme described in Section~\ref{ssec:RLNC}.

To achieve a desired average coding density in the coefficients, for each
of them, we utilize a set of Bernoulli random variables all with parameter
$d$ as it success probability, i.e. $\mathbb{B}_j \sim Bernoulli(d)\ ,\ \forall
j \in [1,g]$. In this way, we can represent a coded packet in \ac{SRLNC} as:
%
\begin{align} \label{eq:sparse_coded_packet}
    C_i  &= \bigoplus_{j=1}^{g} \mathbb{B}_j v_{ij} \otimes P_j ,\ v_{ij}
    \neq 0\
    \forall i \in [1,\ldots) ,\ \
    d \in
    \begin{cases}
        {(0,0.5]} & ,\ q = 2 \\
        {(0,1]}   & ,\ q > 2
    \end{cases}
\end{align}
%
In~\eqref{eq:sparse_coded_packet}, we have the requirement for the
coding coefficient to not be zero, since we want to ensure a coding
coefficient is generated for any random trial, where
$\mathbb{B}_j = 1$. Therefore, in our implementation of \ac{SRLNC} we
exclude the zero element and then pick uniformly distributed random
elements from $GF(q)-\{ 0 \}$. Also, we have specified a dependency on
the field size for practical density values. In case of employing
$GF(2)$, the maximum plausible density is restricted up to 0.5 since
higher values incur in more frequent \ac{l.d.} coded
packets~\cite{realworld_nc2013}.

@@

The drawback of this scheme is that coded packets from the encoder
become more \ac{l.d.} on each other as the density is reduced.  This
leads to transmission overhead since another coded packet is required
to be sent for every reception of a redundant packet.  Also, this
method may still generate a coded packet, which does not contain any
information. For example, we might find the case where
$\left[\mathbb{B}_1, \ldots, \mathbb{B}_g \right] = \mathbf{0}$ may
happen frequently for low densities. In that case, the encoder
discards the coded packet and tries to generate a new one to avoid the
negative impact on overall system performance.

\subsubsection{Method 2: Sampling the Amount of Packets to Combine}
The method described from \eqref{eq:sparse_coded_packet} results
in a fast implementation in terms of execution time for
$d \geq 0.3$~\cite{practicalview_tsnc2015}. Nevertheless, for $d < 0.3$
the implementation becomes slow given that more Bernoulli trials are
required in total for all random variables to generate a coded packet.
Thus, we introduce a second method that permits a fast implementation
for low coding densities~\cite{practicalview_tsnc2015}.

For this method, we first obtain the amount of packets
that we will combine to create a coded packet. To do so, a random
number, $\mathbb{N}$, of the original packets is used to produce a
coded packet. For our case, $\mathbb{N}$ is binomially distributed
with parameters $g$ for the number of trials and $d$ for its success
probability, e.g. $\mathbb{N} \sim Binomial(g,d)$. However,
when sampling from this distribution, the case
$\mathbb{N} = 0$ occurs with a non-zero probability. In order to handle
this special case, our implementation considers
$\mathbb{K} = \max(1,\mathbb{N})$ as the final amount of packets to be
mixed together. In this way, we always ensure that at least one packet is
encoded.

The only caveat is that the case of $\mathbb{K} = 1$ occurs slightly
more often than $\mathbb{N} = 1$ in the original distribution, but for
the considered density range in this method, this is not a significant
modification~\cite{practicalview_tsnc2015}. Then, once the
distribution for the number of packets to mix has been defined, we
sample a value $n$ from $\mathbb{N}$ and compute $k =
\max(1,n)$. Later, we create a create a set $\mathcal{K}$ with
cardinality $k$, e.g. $|\mathcal{K}| = k$, where the elements in
$\mathcal{K}$ are the indexes of the packets that are going to be
considered for making a coded packet. To compute the indexes
of the set $\mathcal{K}$, we do the following algorithm in pseudo-code: \\
\ \\
\begin{algorithm}[H]
 \label{alg:k_set}
 \KwData{$k$: Size of $\mathcal{K}$. $g$: Generation Size}
 \KwResult{$\mathcal{K}$: The set of non-repeated indexes}
 $\mathcal{K}$ = \{ \}\;
 \While{$|\mathcal{K}| \neq k$ }{
  $i$ = Sample from $\mathbb{U}(1,g)$\;
  \If{$i \notin \mathcal{K}$}{
   Insert $i$ in $\mathcal{K}$\;
   }
 }
 \caption{Computation of the set of indexes for packet combination in SRLNC.}
\end{algorithm}
\ \\
In Algorithm~\ref{alg:k_set}, the notation $\mathbb{U}(1,g)$ stands
for a uniform discrete random variable with limits $1$ and $g$. The
pseudo-code in Algorithm~\ref{alg:k_set} indicates that the set $\mathbb{K}$
is filled with non-repeated elements taken uniformly at random from
$[1,\ldots,g]$. Finally, once the set of indexes has been defined, a coded
packet using \ac{SRLNC} can be written as shown in
\eqref{eq:sparse_coded_packet2}:
%
\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{m \in \mathcal{K}} v_{im} \otimes P_{m},\ v_{im} \neq 0,\ \forall i \in [1,\ldots)
    %L \subseteq \{1,\dots,g\}: |L|=k ,\
\end{align}

@@

Therefore, in \ac{TSNC}, a simple procedure for controlling this probability
is to gradually increase the coding density as the \ac{dof}\footnote{We refer
as degrees of freedom to the dimension of the linear span from the coded
packets that a decoder has at a given point} increases, e.g., as the
cardinality of the set of \ac{l.i.} coded packets increases. This enables
\ac{TSNC} to significantly reduce the complexity, particularly in the
beginning of the transmission process and also to control the decoding
delay. For \ac{TSNC}, we define the budget, $b \geq g$, to be a target
number of coded packets a transmitter wants to send to a receiver
for decoding a generation of $g$ packets. In some scenarios, we may set
$e$ defined as the excess of \ac{l.d.} packets sent from the encoder.
This helps to also define the budget as $b = g + e$.

The difference between $b$ and $g$ is equal to the losses in the channel and the
losses due to linear dependencies. Therefore, in a lossless channel, the budget
is
\begin{equation}\label{eq:budget}
    b(g,d)=\sum^{g-1}_{i=0}\frac{1}{P(i,g,d)},
\end{equation}
where $P(i,g,d)$ is the probability of receiving an innovative coded packet
after receiving $i$ \ac{l.i.} packets with a coding density $d$. In our implementations, we considered the the upper bound for the innovation probability from \cite{feizi2014tunable}, given as
\begin{equation}
    P(i,g,d)\geq 1-(1-d)^{g-i}
\end{equation}

Provided a desired budget $b$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k),\
k \in [1, \ldots)$ as the rank $r$ of the decoder\footnote{The rank of
a decoder is the rank of its coding matrix at a given point in the
transmissions process.} where a decoder has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in [1,\ldots)
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in~\cite{practicalview_tsnc2015}.

The transmissions made between feedback packets are called the density
regions since the same estimated density is used for encoding sparse
packets. Once a new feedback is received, an encoder recalculates the
coding density for the new region until the next feedback packet is
received. before the coding density can be estimated, it is essential
that the encoder has a good estimate of the decoder \ac{dof}, but also
the remainder of the total budget. To calculate the density, we use
bisection to estimate a fixed density for the density region that
satisfies the budget for the $k$-th density region:
\begin{equation}
b(s(k-1),r(k),g,d) = \sum^{r(k)}_{i=r(k-1)} \frac{1}{P(i,g,d)} = \sum^{r(k)}_{i=r(k-1)} \frac{1}{1-(1-d)^{g-i}} = \frac{b}{2^k}, k\in [1,...).
\end{equation}
%
% Where the budget is half of the remaining budget for all the density regions
% except the last one, to which it is assigned the remaining of the budget.
The feedback scheme based on the rank reports
of~\eqref{eq:feedback_dofs} roughly splits the remaining coding
regions in two halves. In other words, the first half is the region
where the encoder currently operates. Once an encoder finishes with
this region, it proceeds with second half where again it splits this
into to two region and so on until it finishes on transmitting the
whole generation. This is also the case for the budget that is split
equally among the two new regions. The very last region will be
assigned the full remainder of the total budget and the coding density
will not vary.

@@

When designing multi-threaded algorithms for network coding
operations, it is possible to implement the decoding by combining the
matrix inversion and the matrix multiplication, e.g., performing the
Gauss-Jordan algorithm over the coding matrix $\textbf{V}$ while
performing, in parallel, row operations on the coded data
$\textbf{C}$. For example, in~\cite{5061951} and~\cite{4262451} the
parallelization of the row operations are optimized for \ac{GPU} and
\ac{SMP} systems respectively. However, the parallelization of such
operations provide limited speeds up for small block sizes
($\leq 2048$ bytes). The reason is that to operate in a parallel
fashion over the same coded packet $C_i$ requires strained
synchronization.
